---
title: "Assignment 1"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

## Preliminaries

We're using the [Tidyverse](https://www.tidyverse.org/).

```{r}
library(tidyverse)
```

## Load the survey data

```{r}
d <- read_csv("survey-2022-1.csv")
```

```{r}
summary(d)
```

## Analysis

```{r}
ggplot(data=d,aes(x=Belts)) + 
  geom_histogram() +
  labs(title="Distribution of number of belts",
       caption="10 was the maximum allowable value")
```

---

MKT382 Data Analysis & Visualization
Prof Jamie Stephens
Yige Ren 
Sept 6, 2022


## Graph1: Descriptive Statistics of Continent MSM Students are from
Since Continent is a categorical variable, using bar plot can depict the distribution well:

```{r}
ggplot(data=d,aes(x=Continent)) + 
  geom_bar(width=0.5,fill="pink") +
  labs(title="Distribution of Continent", caption="Source: MSM Survey")+
  geom_text(stat="count",aes(label=..count..))
```

Observation: This graph shows that more than half of MSM students come from North America, which is the dominant continent; the second common continent students are from is Asia. There's only a few students come from Africa&South America.



---

## Graph2: Relationship between the Degree Students Like Math and Their Undergraduate Major

To show the range of students' ratings, I chose to use box plot:


```{r}
ggplot(data=d,aes(x=Undergrad, y=LikeMath)) + 
  geom_boxplot(width=0.5) +
  labs(title="Undergrad Major & Degree of Liking Math", x="Undergrad Major", y="Degree of Liking Math", caption="Source: MSM Survey")

```

Observation: From the graph, it can be seen that students with a business or engineering major relatively have a higher level of interests in math. Students with a liberal arts degree and other degrees have lower liking in math. Among all the majors, Communication major students have the highest variation in math interests.



---


## Graph3: Scatter Plot of Computer Type & Tablet Type
I want to learn whether students with Mac tend to have a pad compared to Windows users, and also see what's the distribution of Phone type data on this problem.

```{r}
ggplot(d, aes(x=Computer, y=Tablet, color=Phone))+
  geom_point()+
  geom_jitter(width=0.15)+
  labs(title="Computer Type & Tablet Type", caption="Source: MSM Survey")
  
```  

Observation: From the plot, there are more data points in the Mac&iPad region compared to Windows&iPad region, which can be inferred that students do tend to buy an iPad if they are Mac users or vise versa. And among all MSM students, there's only one student using Android. 



---

## Graph4: Analyze Whether Students Have R Experience in Terms of Their Type of Computer 

To show the distinction of R experience by looking at their type of computer, I chose to employ a bar plot to show the categorical variable Computer, and use different colors filling it to show different R experience.

```{r}
ggplot(data=d,aes(x=Computer, fill=R)) + 
  geom_bar(width=0.5) +
  labs(title="R Experience by Computer Type", x="Computer Typee", caption="Source: MSM Survey")+
  geom_text(stat = "count", aes(label=..count..),position="stack", size=4)
```

Observation: Most of MSM students use MAC, among which around 1/3 students don't have R experiences; while in Windows users, the amount of students with and without R experience is roughly the same.




---

## JustforFun1: Are Asian students really like math more????

I want to check about the long existing "stereotype" that Asian students love math and they are good at math. So I group MSM students with different continents they are from, and calculate the mean of their LikeMath rating based on that. I chose to use a bar plot to show the result.

```{r}
d%>%
  group_by(Continent)%>%
  summarise(mean=mean(LikeMath, na.rm=TRUE)) %>%
  ggplot(aes(x=Continent, y=mean))+
  geom_col()+
  labs(title="LikeMath by Continent", y="Mean of LikeMath", caption="Source: MSM Survey")+
  geom_text(aes(label=round(mean,2)), color="white",position="fill")
```

Observation: According to the result, the mean of LikeMath is indeed higher in Asian & South America students, but Asian students is a little bit lower than the South American group. Considering there are only two students who are from SA and one of them like math a lot (rate 8), and there are 16 students coming from Asia, I think such differences might be "biased" in leading to conclusions. Thus in this tiny sample, I'm afraid that the long-held stereotype might be held even longer :)



---


## JustforFun2: Relationship between R Experience & Python Experience

I want to learn whether there's correlation between students' R experiences and Python experiences, so I plan to use a scatter plot to observe.

To begin with, I will transfer the Boolean variables to 0&1.

```{r}
d %>%
  mutate(R_Experience=
           ifelse(R=="Yes", 1, 0), 
         Python_Experience=
           ifelse(Python=="Yes", 1, 0)) %>%
  ggplot(aes(R_Experience, Python_Experience))+
  geom_point()+
  geom_jitter(width=0.3)+
  geom_smooth(method="glm", se=FALSE, method.args=list(family=binomial))+
  labs(title="Simple Regression on R Experience&Python Experience", caption="Source: MSM Survey")
```

Observation: From the graph, it seems that students who have R experience also tends to have Python Experience; but such correlation is weak. Also the model I deployed has limitations, but logit model is the most close one I can come up with in helping me solve a logic-logic relationship right now...according to Prof.Sager, there's no perfect models, good enough is good enough...But I will definitely stay tuned and hope to grab some better model & simulations from future classes :)


```{r}
d%>% count(Undergrad) %>% ggplot(aes(x=Undergrad,y=n))+
  geom_bar(stat = "identity")+
  coord_flip()

d%>% count(Undergrad) %>% ggplot(aes(x=reorder(Undergrad,n),y=n))+
  geom_bar(stat = "identity")+
  coord_flip()

d%>% count(Undergrad) %>% ggplot(aes(x=Undergrad,y=n))+
  geom_bar(stat = "identity")+
  coord_flip()
```